---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wespr

<!-- badges: start -->
[![R-CMD-check](https://github.com/BCWF-Wetlands/wespr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/BCWF-Wetlands/wespr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The B.C. Wildlife Federation’s Wetlands Workforce project is a collaboration with conservation organizations and First Nations working to maintain and monitor wetlands across British Columbia. https://bcwf.bc.ca/initiatives/wetlands-workforce/.  

The [Wetland Ecosystem Services Protocol](https://wetlandsworkforce.ca/portfolio/wetlands-ecosystem-services-protocol/) (WESP) model is a standardized method for assessing the function and value of the services provided by wetlands.

wespr is an R package to help automate the validation of WESP data and to calculate 
WESP indicators in a reproducible manner.

## Installation

You can install the development version of wespr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("BCWF-Wetlands/wespr")
```

## Example usage

```{r example}
library(wespr)
```

Once the package is loaded, we can read in the WESP data from a csv, which should be in a
standardized format as output from the scripts at https://github.com/BCWF-Wetlands/WESP_Calculator.

An example format of the file is shown, with the first 10 questions (rows) and 6 sites (columns) shown:

```
Question,1,2,3,4,5,6
F1_1,1,0,1,1,1,2
F1_2,0,3,3,2,0,2
F1_3,1,0,3,1,1,1
F1_4,1,4,1,1,1,3
F1_5,1,0,1,1,1,0
F1_6,1,1,1,1,1,4
F2_A1,0,0,0,0,0,1
F2_A2,0,0,0,0,0,0
F2_B1,0,1,0,0,0,0
F2_B2,1,0,1,1,1,0
```

We have included an example file in the package, which we can use for demonstration purposes:

```{r}
wesp_file <- system.file("input_data/wetFlat.csv", package = "wespr")
wesp_data <- load_wesp_data(wesp_file)

head(wesp_data)
```

Next we convert the data into a special `wesp_site` object. This process validates
the input data, and will give informative errors if a question is not answered correctly.
It also calculates many "derived values" from the questions, which are common inputs into different
indicator calculations.

If there is more than one site in the input data, you can select which site you
would like to use --- if no site is selected it will default to choosing the first 
site in the file.

```{r}
site <- as.wesp_site(wesp_data, site = 1)
```

You should not normally need to work with the internals of the `wesp_site` object
as it is quite complex, however we can get an overview of what is in it by just 
typing the name of the object:

```{r}
site
```

We can see that the responses are loaded, with a few flagged as incomplete, and
the derived values have been calculated. However, the indicators have not yet
been calculated.

Now, we can calculate the indicator scores with the `calc_indicators()` function.

You must assign the output of this function back to the original `wesp_site` 
object, as it updates the `indicators` with the calculated values:

```{r}
site <- calc_indicators(site)
```

If we view the site object again, we can see the calculated indicator scores.

```{r}
site
```

We probably want to get the indicator scores out as a usable object. We can do 
that with `get_indicator_scores()`, which gives them to us as a `data.frame`:

```{r}
ind_scores <- get_indicator_scores(site)

ind_scores
```

Similarly, we can extract the original responses with `get_responses()`:

```{r}
get_responses(site)
```

We can also get out a data.frame of derived values, those values which are 
calculated from the responses, and used as inputs into many indicators.

```{r}
get_derived_values(site)
```

## Calculating multiple site scores

In most cases we wish to calculate indicator scores for multiple site. This can 
be done with the `calculate_multi_site()` function. Which requires the data to 
firstly be loaded using the `load_wesp_data()` function. We can then iterate 
through all sites.


```{r}
calculate_multi_site(wesp_data)

```



## Calculating Jenks scores






## Anatomy of the `wesp_site` object

As mentioned above, this is a large, complex object, and it is almost always 
best to interact with the `wesp_site` object via the provided functions. If 
there is functionality that is missing please 
[open an issue](https://github.com/BCWF-Wetlands/wespr/issues). 

It has four elements:

```{r}
length(site)
names(site)
```

The first three (`site_name`, `questions`, and `derived_values`) are populated when 
the `wesp_site` object is created, but the `indicators` element is just a placeholder, until `calc_indicators()` has been run:

```{r}
site$site_name

# The questions element is large and complex, so we will just show the names:
names(site$questions)

site$derived_values

site$indicators
```

## Development

### Updating internal data

The wespr package contains two internal datasets which facilitate question validation and indicator calculations: `question_metadata` and `indicator_weightings`.

These are updated by:

1. Update the Google Sheet at:  https://docs.google.com/spreadsheets/d/1l2h7Z65H5z0cKv_gvorxkT6k9LC7ZKLS/edit#gid=924841838 (you need to be authorized to edit this document)

2. Run the script in `data-raw/internal-data.R`

3. Reload/reinstall the package.

4. Run `devtools::check()` to make sure the changes you made to the Google Sheet haven't broken any existing behaviour.

5. Commit the changes to `R/sysdata.rda` (this file contains both datasets)


