---
title: "WESP site report"
output: word_document
params:
  calibration_scores_eco: calibration_scores_eco
  calibration_scores_summary: calibration_scores_summary,
  classed_df: classed_df
  output_dir: output_dir
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(wespr)
library(dplyr)
library(ggplot2)
library(sf)
library(stringr)

#testing lines 
calibration_scores_eco = calibration_scores_eco # calibration dataset
calibration_scores_summary = calibration_scores_summary # summary of cal set for high and lows - used in graphics
classed_df = classed_df # results per site 
output_dir = output_dir # output dir
# end testing line 

```


```{r cars, echo = FALSE}
#summary(cars)
ecop <- unique(calibration_scores_summary$ecoprovince)
sname <- unique(classed_df$site)

# add map to show where in the province

```

**Report generated on :** `r Sys.Date()`

**Site Name :** `r sname`

**Ecoprovince :** `r ecop`


## 2. Site Location 

```{r generate map, echo = FALSE}
if (ecop == "GD"){
  ecop = "GED"

  # need to add some fixes for TBP here 
  }
  
ecoi <- ecoprovince_sp |>
  dplyr::filter(ECOPROVINCE_CODE == ecop)

map1 <- ggplot() +
  geom_sf(data = ecoprovince_sp$geometry, fill = "lightgrey") +
  geom_sf(data = ecoi, fill = "lightblue") +
  labs(title = paste0("Ecoprovince : ", stringr::str_to_title(unique(ecoi$ECOPROVINCE_NAME))))+
  theme_void()

```

```{r plot map, echo = FALSE, fig.width=3.5, fig.height=3.25}
map1

```

## Wetland Scores


```{r, echo = FALSE}
# generate the background information for Calibration thresholds
# Calculate the numbers per group or by plot

cal_no <- calibration_scores_eco

wcols <- names(calibration_scores_eco)
## remove the site column
wcols <- wcols[!wcols %in% c("site", "wetland_id", "ecoprovince")]
wcols <- unique(sub("^([^_]*_[^_]*).*", "\\1", wcols))

# loop through the data and get the min and max values of the raw scores

outsum <- purrr::map(wcols, function(x) {
  # get the columns for each service
  # x <- wcols[1]

  tw <- calibration_scores_eco |>
    dplyr::select(dplyr::starts_with(x)) |>
    dplyr::select(-dplyr::ends_with("_norm"))
  names(tw) <- c("jenks", "raw")

  tww <- tw |>
    # dplyr::group_by(jenks) |>
    # dplyr::summarise(n = n(),
    #                  min = min(raw),
    #                  max = max(raw)) |>
    dplyr::mutate(
      service = x,
      service_name = unique(sub("^([^_]*).*", "\\1", service)),
      service_type = unique(sub("^[^_]*_", "", service))
    )
  tww
}) |> dplyr::bind_rows()

outsum <- outsum |>
  dplyr::mutate(jenks = ordered(jenks, levels = c("H", "M", "L"))) |>
  mutate(service_full_name = case_when(
    service_name == "AM" ~ "Amphibian Habitat (AM)",
    service_name == "APP" ~ "Aquatic Primary Productivity (APP)",
    service_name == "CP" ~ "Carbon Preservation (CP)",
    service_name == "CRI" ~ "Cultural Recreational Importance (CRI)",
    service_name == "FH" ~ "Fish Habitat (FH)",
    service_name == "FR" ~ "Fire Resistance (FR)",
    service_name == "KMH" ~ "Keystone Mammal Habitat (KMH)",
    service_name == "NR" ~ "Nitrate Removal and Retention (NR)",
    service_name == "OE" ~ "Organic Matter Export (OE)",
    service_name == "PD" ~ "Native Plant Diversity (PD)",
    service_name == "POL" ~ "Pollinator Habitat (POL)",
    service_name == "PR" ~ "Phosphorus Retention (PR)",
    service_name == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
    service_name == "SENS" ~ "Wetland Sensitivity (SENS)",
    service_name == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
    service_name == "SR" ~ "Sediment Retention and stabilization (SR)",
    service_name == "STR" ~ "Wetland Stressors (STR)",
    service_name == "WB" ~ "Waterbird Habitat (WB)",
    service_name == "WS" ~ "Water Storage and Delay (WS)"
  ))

# Calculate the numbers per group or by plot
cal_no <- calibration_scores_eco

# grouped scores
outsum_f <- outsum |>
  dplyr::filter(service_type == "f")


```


NOTE: A score of 0 does not always mean the function or value is absent from the wetland. It usually
means that this wetland has equal or less capacity than the lowest-scoring one, for that function or
value, from among the 100+ calibration wetlands that were assessed previously for the EcoProvince.


#### Output Descriptions

The **Score** columns present the numeric score of a function, value, or attribute after it has been
mathematically adjusted (standardized) to a full 0-10 point scale, based on calibration data from 100+
test sites.

The **Rating** columns indicate which of three rating categories [Lower (L), Moderate (M), Higher(H)]
each normalized score is assigned to. Ratings convey the relative meaning of the numeric score and
allow for comparison across different functions and benefits. The breaks between ratings differ for
each function and benefit, based on the distribution of scores from the 100+ calibration wetlands.
<br>



```{r, echo = FALSE, include = FALSE}
# # generate raw values plots based on themes 
# 
# # functions 
# ffraw <- classed_df |> 
#   mutate(service_name = indicator) |> 
#   mutate(service_full_name = case_when(
#     service_name == "AM" ~ "Amphibian Habitat (AM)",
#     service_name == "APP" ~ "Aquatic Primary Productivity (APP)",
#     service_name == "CP" ~ "Carbon Preservation (CP)",
#     service_name == "CRI" ~ "Cultural Recreational Importance (CRI)",
#     service_name == "FH" ~ "Fish Habitat (FH)",
#     service_name == "FR" ~ "Fire Resistance (FR)",
#     service_name == "KMH" ~ "Keystone Mammal Habitat (KMH)",
#     service_name == "NR" ~ "Nitrate Removal and Retention (NR)",
#     service_name == "OE" ~ "Organic Matter Export (OE)",
#     service_name == "PD" ~ "Native Plant Diversity (PD)",
#     service_name == "POL" ~ "Pollinator Habitat (POL)",
#     service_name == "PR" ~ "Phosphorus Retention (PR)",
#     service_name == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
#     service_name == "SENS" ~ "Wetland Sensitivity (SENS)",
#     service_name == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
#     service_name == "SR" ~ "Sediment Retention and stabilization (SR)",
#     service_name == "STR" ~ "Wetland Stressors (STR)",
#     service_name == "WB" ~ "Waterbird Habitat (WB)",
#     service_name == "WS" ~ "Water Storage and Delay (WS)"
#   ))
# 
# ffraw <- ffraw |> 
#   select(service_type, value, service_full_name) |> 
#   mutate(service_type_full = case_when(
#     service_type == "f" ~ "function",
#     service_type == "b" ~ "benefit"
#   ))
# 
# ffraw <- ffraw |> 
#   mutate(colour = case_when(
#       value > -1 & value < 1 ~ "#FDE725FF",
#       value > 1 & value < 4 ~ "#20A387FF",
#       value < -4 ~ "#482677FF",
#       value < -1 ~ "#20A387FF",
#       value > 4 ~ "#482677FF"))
# 
# 
# raw_plot <- ggplot(ffraw, aes(x = service_type_full, y = service_full_name)) +
#   geom_tile(aes(fill = value)) +
#   geom_text(aes(label = round(value, 1))) +
#   #scale_fill_continuous("blue")
#   theme(palette.colour.continuous = c("#FDE725FF", "#20A387FF", "#453781FF"))+
#   #theme(palette.colour.discrete = c("#FDE725FF", "#20A387FF", "#482677FF"))+
#   labs(
#     x = "",
#     y = "Ecosystem Function"
#   )+
#   theme(legend.position="none")


```


NOTE: A score of 0 does not always mean the function or value is absent from the wetland. It usually
means that this wetland has equal or less capacity than the lowest-scoring one, for that function or
value, from among the 100+ calibration wetlands that were assessed previously for the EcoProvince.


#### Output Descriptions

The **Score** columns present the numeric score of a function, value, or attribute after it has been
mathematically adjusted (standardized) to a full 0-10 point scale, based on calibration data from 100+
test sites.

The **Rating** columns indicate which of three rating categories [Lower (L), Moderate (M), Higher(H)]
each normalized score is assigned to. Ratings convey the relative meaning of the numeric score and
allow for comparison across different functions and benefits. The breaks between ratings differ for
each function and benefit, based on the distribution of scores from the 100+ calibration wetlands.
<br>


```{r raw_plots, echo = FALSE, width = 10}
raw_plot

```


### 2.2 Ecosystem Function Scores


```{r, echo = FALSE}
# # generate raw values plots based on themes 
# 
# # functions 
# 
# ff <- classed_df |> 
#   mutate(service_name = indicator) |> 
#   mutate(service_full_name = case_when(
#     service_name == "AM" ~ "Amphibian Habitat (AM)",
#     service_name == "APP" ~ "Aquatic Primary Productivity (APP)",
#     service_name == "CP" ~ "Carbon Preservation (CP)",
#     service_name == "CRI" ~ "Cultural Recreational Importance (CRI)",
#     service_name == "FH" ~ "Fish Habitat (FH)",
#     service_name == "FR" ~ "Fire Resistance (FR)",
#     service_name == "KMH" ~ "Keystone Mammal Habitat (KMH)",
#     service_name == "NR" ~ "Nitrate Removal and Retention (NR)",
#     service_name == "OE" ~ "Organic Matter Export (OE)",
#     service_name == "PD" ~ "Native Plant Diversity (PD)",
#     service_name == "POL" ~ "Pollinator Habitat (POL)",
#     service_name == "PR" ~ "Phosphorus Retention (PR)",
#     service_name == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
#     service_name == "SENS" ~ "Wetland Sensitivity (SENS)",
#     service_name == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
#     service_name == "SR" ~ "Sediment Retention and stabilization (SR)",
#     service_name == "STR" ~ "Wetland Stressors (STR)",
#     service_name == "WB" ~ "Waterbird Habitat (WB)",
#     service_name == "WS" ~ "Water Storage and Delay (WS)"
#   )) |> 
#   select( service_type, value, service_name, service_full_name) |> 
#   mutate(service_type_full = case_when(
#     service_type == "f" ~ "function",
#     service_type == "b" ~ "benefit"
#   ))
# 
# ggplot(ff, aes(x = service_type_full, y = service_full_name)) +
#   geom_tile(aes(fill = value)) +
#   geom_text(aes(label = round(value, 1))) +
#   #scale_fill_continuous("blue")
#   theme(palette.colour.continuous = c("#DEEBF7", "#9ECAE1", "#3182BD"))+
#   labs(
#     x = "Ecosystem Score",
#     y = "Ecosystem Function"
#   )

```



```{r overall plots, echo=FALSE, fig.width=8, fig.height=7}
tt <- calibration_scores_summary |>
  group_by(.data$service) |>
  select(-n) |>
  tidyr::pivot_wider(names_from = jenks, values_from = c(min, max)) |>
  dplyr::mutate(
    H = 10.1 - min_H,
    M = min_H - min_M,
    L = min_M
  )

ttt <- tt |>
  tidyr::pivot_longer(
    cols = c(H, M, L, -min_H, -min_L, -min_M, -max_H, -max_L, -max_M),
    names_to = "catergory",
    values_to = "threshold"
  ) |>
  dplyr::select(-c(min_H, min_L, min_M, max_H, max_L, max_M))

ttt <- ttt |>
  dplyr::mutate(class = ordered(catergory, levels = c("H", "M", "L"))) |>
  mutate(service_full_name = case_when(
    service_name == "AM" ~ "Amphibian Habitat (AM)",
    service_name == "APP" ~ "Aquatic Primary Productivity (APP)",
    service_name == "CP" ~ "Carbon Preservation (CP)",
    service_name == "CRI" ~ "Cultural Recreational Importance (CRI)",
    service_name == "FH" ~ "Fish Habitat (FH)",
    service_name == "FR" ~ "Fire Resistance (FR)",
    service_name == "KMH" ~ "Keystone Mammal Habitat (KMH)",
    service_name == "NR" ~ "Nitrate Removal and Retention (NR)",
    service_name == "OE" ~ "Organic Matter Export (OE)",
    service_name == "PD" ~ "Native Plant Diversity (PD)",
    service_name == "POL" ~ "Pollinator Habitat (POL)",
    service_name == "PR" ~ "Phosphorus Retention (PR)",
    service_name == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
    service_name == "SENS" ~ "Wetland Sensitivity (SENS)",
    service_name == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
    service_name == "SR" ~ "Sediment Retention and stabilization (SR)",
    service_name == "STR" ~ "Wetland Stressors (STR)",
    service_name == "WB" ~ "Waterbird Habitat (WB)",
    service_name == "WS" ~ "Water Storage and Delay (WS)"
  ))

# match format for site data
classed_df <- classed_df |>
  dplyr::mutate(
    service = paste0(indicator, "_", service_type),
    service_name = indicator,
    threshold = value
  ) |>
  mutate(service_full_name = case_when(
    service_name == "AM" ~ "Amphibian Habitat (AM)",
    service_name == "APP" ~ "Aquatic Primary Productivity (APP)",
    service_name == "CP" ~ "Carbon Preservation (CP)",
    service_name == "CRI" ~ "Cultural Recreational Importance (CRI)",
    service_name == "FH" ~ "Fish Habitat (FH)",
    service_name == "FR" ~ "Fire Resistance (FR)",
    service_name == "KMH" ~ "Keystone Mammal Habitat (KMH)",
    service_name == "NR" ~ "Nitrate Removal and Retention (NR)",
    service_name == "OE" ~ "Organic Matter Export (OE)",
    service_name == "PD" ~ "Native Plant Diversity (PD)",
    service_name == "POL" ~ "Pollinator Habitat (POL)",
    service_name == "PR" ~ "Phosphorus Retention (PR)",
    service_name == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
    service_name == "SENS" ~ "Wetland Sensitivity (SENS)",
    service_name == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
    service_name == "SR" ~ "Sediment Retention and stabilization (SR)",
    service_name == "STR" ~ "Wetland Stressors (STR)",
    service_name == "WB" ~ "Waterbird Habitat (WB)",
    service_name == "WS" ~ "Water Storage and Delay (WS)"
  ))


# summary of the ranks overall
all_ranks_f <- classed_df |>
  filter(service_type == "f") |>
  group_by(calibration_scores_summary) |>
  summarise(n = n())

# functions
ft <- ttt |>
  dplyr::filter(service_type == "f")

fclass <- classed_df |>
  dplyr::filter(service_type == "f") |>
  filter(!is.na(value)) |>
  left_join(outsum)


fn_plot <- ggplot(ft, aes(x = threshold, y = service_full_name)) +
  geom_bar(aes(fill = class), stat = "identity", position = "stack") +
  scale_fill_viridis_d() +
  # geom_text(data= fclass,aes(x= max,y=service_full_name,label=n), color = "darkgrey",vjust=0) +
  geom_point(data = fclass, aes(x = threshold, y = service_full_name, size = 1.25), colour = "darkgrey") +
  labs(
    x = "Ecosystem Score",
    y = "Ecosystem Function"
  ) +
  scale_size(guide = 'none')+
  theme_minimal()

```

This site ecosystem function scores included `r all_ranks_f |> filter(calibration_scores_summary == "H") |> pull(n)` High ranked functions, `r all_ranks_f |> filter(calibration_scores_summary == "M") |> pull(n)` Medium ranked functions and `r all_ranks_f |> filter(calibration_scores_summary == "L") |> pull(n)` Low ranked functions. 


The plot below shows where `r sname` scores relative to the baseline values for the `r ecop` Ecoprovince. The grey dot indicated the score for the single site, while the yellow, aqua and purple bars show the threshold values for Low, Medium, High ranked sites within the Ecoprovince.  

```{r, echo = FALSE, fig.width=8, fig.height=6}
 fn_plot
```
<br>

#### High Ranked Functions

```{r, echo = FALSE}
fclass <-  classed_df|> 
   dplyr::filter(service_type == "f") |> 
   filter(!is.na(value)) |> 
   select(service_full_name, calibration_scores_summary,service_name )

```

Ecosystems for the high ranked functions include: `r fclass |> filter(calibration_scores_summary == "H") |> pull(service_full_name) |> paste(collapse = ", ")`. 
<br>

#### Low Ranked Functions

Ecosystems for the low ranked functions include: `r fclass |> filter(calibration_scores_summary == "L") |> pull(service_full_name) |> paste(collapse = ", ")`. 


<br>

### 2.3 Ecosystem Benefit Scores 

```{r, echo = FALSE, fig.width=8, fig.height=6}
 # Benefit
 bt <- ttt |> 
   dplyr::filter(service_type == "b") 
 
 bclass <-  classed_df|> 
   dplyr::filter(service_type == "b")|> 
   filter(!is.na(value))
 
 
bn_plot <-  ggplot(bt , aes(x = threshold,y = service_full_name  )) +
    geom_bar(aes(fill = class), stat = "identity", position = "stack")+
    scale_fill_viridis_d()+ 
   #  geom_text(data= dfTab,aes(x=x,y=Freq,label=lab),vjust=0) +
    geom_point(data =  bclass, aes( x = threshold, y = service_full_name, size = 1.25 ), colour = "darkgrey")+
   labs(
        x = "Ecosystem Score",
        y = "Ecosystem Benefit") +
   scale_size(guide = 'none')+
   theme_minimal()

 #summary of the ranks overall 
 all_ranks_b <- bclass |> 
  group_by(calibration_scores_summary) |>
  summarise(n = n())
   
```

This site ecosystem benefit scores included `r all_ranks_b |> filter(calibration_scores_summary == "H") |> pull(n)` High ranked benefits, `r all_ranks_b |> filter(calibration_scores_summary == "M") |> pull(n)` Medium ranked benefits and `r all_ranks_b |> filter(calibration_scores_summary == "L") |> pull(n)` Low ranked benefits 

```{r echo = FALSE, fig.width=8, fig.height=6}
 bn_plot 
```

<br>

#### High Ranked Benefits

```{r, echo = FALSE}
bclass <-  classed_df|> 
   dplyr::filter(service_type == "b") |> 
   filter(!is.na(value)) |> 
   select(service_full_name, calibration_scores_summary,service_name)

```

Ecosystems for the high ranked benefits include: `r bclass |> filter(calibration_scores_summary == "H") |> pull(service_full_name) |> paste(collapse = ", ")`. 
<br>

#### Low Ranked Benefits
Ecosystems for the high ranked benefits include: `r bclass |> filter(calibration_scores_summary == "L") |> pull(service_full_name) |> paste(collapse = ", ")`.


```{r, eval = FALSE, echo = FALSE}

# Still to add

### review all indicators? or review which ones are producing high results? 

### Individual sensitivity analysis? I think this will take too long? 

```

<br>

<br>
