---
title: "WESP Site Report"
output: word_document
params:
  calibration_scores_eco: calibration_scores_eco
  calibration_scores_summary: calibration_scores_summary,
  EcoP_calibration: EcoP_calibration,
  ecoprovince_sp: ecoprovince_sp,
  classed_df: classed_df
  output_dir: output_dir
  x_loc: x_loc
  y_loc: y_loc
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(wespr)
library(dplyr)
library(ggplot2)
library(sf)
library(stringr)
library(gt)

# # # #testing lines 
# calibration_scores_eco = calibration_scores_eco # calibration dataset
# calibration_scores_summary = calibration_scores_summary # summary of cal set for high and lows - used in graphics
# classed_df = classed_df # results per site
# output_dir = output_dir # output dir
# x_loc =  x_loc
# y_loc = y_loc
# EcoP_calibration = EcoP_calibration
# 
## end testing line 

```


```{r define names, echo = FALSE}
#summary(cars)
ecop <- unique(calibration_scores_summary$ecoprovince)
sname <- unique(classed_df$site)
wlname <- unique(classed_df$wetland_id)
ecop_cal <- ifelse(is.na(EcoP_calibration),ecop, EcoP_calibration)


```

**Report generated on :** `r Sys.Date()`

**Site Name :** `r sname`

**Wetland Id :** `r wlname`

**Wetland location X:** `r x_loc` 

**Wetland location Y:** `r y_loc`

**Wetland Ecoprovince :** `r ecop`

**Calibration Ecoprovince :** `r ecop_cal`

<br>
<br>

## 1. Site Location 

```{r generate map, echo = FALSE}

library(bcdata)
ecoprovince_sp <-  bcdata::bcdc_query_geodata('WHSE_TERRESTRIAL_ECOLOGY.ERC_ECOPROVINCES_SP') |>
  bcdata::collect() |>
  dplyr::select(.data$ECOPROVINCE_CODE, .data$ECOPROVINCE_NAME, .data$geometry) |>
  dplyr::filter(!ECOPROVINCE_CODE %in% c("SAL","NEP" )) |>
  dplyr::mutate(eco_code = dplyr::case_when(
    ECOPROVINCE_CODE == "NBM" ~ "NBM",
    ECOPROVINCE_CODE == "TAP" ~ "BTP",
    ECOPROVINCE_CODE == "BOP" ~ "BTP",
    ECOPROVINCE_CODE == "SBI" ~ "SBI",
    ECOPROVINCE_CODE == "SIM" ~ "SIM",
    ECOPROVINCE_CODE == "SOI"  ~ "SI",
    ECOPROVINCE_CODE == "COM" ~ "CM",
    ECOPROVINCE_CODE == "GED" ~ "GD",
    ECOPROVINCE_CODE == "CEI" ~ "CI"
  ))

if(!is.na(x_loc) & !is.na(y_loc)){
  pts <- sf::st_as_sf(data.frame(x = x_loc, y = y_loc), coords = c("x", "y"), crs = 4326)
  pts <- sf::st_transform(pts, st_crs(ecoprovince_sp))
} 

ecoi <- ecoprovince_sp |>
  dplyr::filter(eco_code == ecop)

map1 <- ggplot() +
  geom_sf(data = ecoprovince_sp$geometry, fill = "lightgrey") +
  geom_sf(data = ecoi, fill = "lightblue") +
  labs(title = paste0("Ecoprovince : ", stringr::str_to_title(unique(ecoi$ECOPROVINCE_NAME))))+
  theme_void()

if(!is.na(x_loc) & !is.na(y_loc)){
  map1 <- map1 + 
    geom_sf(data = pts, color = "red", size = 3) 
}


```

```{r plot map, echo = FALSE, fig.width=3.5, fig.height=3.25}
map1

```
<br>

## 2. Wetland Scores and Ratings

```{r run calibration calculations, echo = FALSE}
# generate the background information for Calibration thresholds
# Calculate the numbers per group or by plot

cal_no <- calibration_scores_eco

wcols <- names(calibration_scores_eco)
## remove the site column
wcols <- wcols[!wcols %in% c("site", "wetland_id", "ecoprovince")]
wcols <- unique(sub("^([^_]*_[^_]*).*", "\\1", wcols))

# loop through the data and get the min and max values of the raw scores

outsum <- purrr::map(wcols, function(x) {
  # get the columns for each service
  # x <- wcols[1]

  tw <- calibration_scores_eco |>
    dplyr::select(dplyr::starts_with(x)) |>
    dplyr::select(-dplyr::ends_with("_norm"))
  names(tw) <- c("jenks", "raw")

  tww <- tw |>
    # dplyr::group_by(jenks) |>
    # dplyr::summarise(n = n(),
    #                  min = min(raw),
    #                  max = max(raw)) |>
    dplyr::mutate(
      service = x,
      service_name = unique(sub("^([^_]*).*", "\\1", service)),
      service_type = unique(sub("^[^_]*_", "", service))
    )
  tww
}) |> dplyr::bind_rows()

outsum <- outsum |>
  dplyr::mutate(jenks = ordered(jenks, levels = c("H", "M", "L"))) |>
   dplyr::mutate(service_full_name =  dplyr::case_when(
    service_name == "AM" ~ "Amphibian Habitat (AM)",
    service_name == "APP" ~ "Aquatic Primary Productivity (APP)",
    service_name == "CP" ~ "Carbon Preservation (CP)",
    service_name == "CRI" ~ "Cultural Recreational Importance (CRI)",
    service_name == "FH" ~ "Fish Habitat (FH)",
    service_name == "FR" ~ "Fire Resistance (FR)",
    service_name == "KMH" ~ "Keystone Mammal Habitat (KMH)",
    service_name == "NR" ~ "Nitrate Removal and Retention (NR)",
    service_name == "OE" ~ "Organic Matter Export (OE)",
    service_name == "PD" ~ "Native Plant Diversity (PD)",
    service_name == "POL" ~ "Pollinator Habitat (POL)",
    service_name == "PR" ~ "Phosphorus Retention (PR)",
    service_name == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
    service_name == "SENS" ~ "Wetland Sensitivity (SENS)",
    service_name == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
    service_name == "SR" ~ "Sediment Retention and stabilization (SR)",
    service_name == "STR" ~ "Wetland Stressors (STR)",
    service_name == "WB" ~ "Waterbird Habitat (WB)",
    service_name == "WS" ~ "Water Storage and Delay (WS)"
  ))

# Calculate the numbers per group or by plot
cal_no <- calibration_scores_eco

# grouped scores
outsum_f <- outsum |>
  dplyr::filter(service_type == "f")

```


```{r build table, echo = FALSE}
# generate the standard table 
lsorder <- c(
  "Hydrologic Group (WSg)",
  "Water Storage and Delay (WS)",
  "Water Quality (WQg)",
  "Sediment Retention and stabilization (SR)",
  "Phosphorus Retention (PR)",
  "Nitrate Removal and Retention (NR)",
  "Aquatic Support (ASg)",
  "Stream Flow and Temperature Support (SFTS)",
  "Organic Matter Export (OE)",
  "Aquatic Primary Productivity (APP)",
  "Aquatic Habitat (AHg)",
  "Fish Habitat (FH)",
  "Waterbird Habitat (WB)",
  "Amphibian Habitat (AM)",
  "Transition Zone Habitat Group (THg)",
  "Keystone Mammal Habitat (KMH)",
  "Raptor and Wetland Songbird Habitat (RSB)",
  "Native Plant Diversity (PD)",
  "Pollinator Habitat (POL)",
  "Other Values or Attributes",
  "Carbon Preservation (CP)",
  "Fire Resistance (FR)",
  "Cultural Recreational Importance (CRI)",
  "Wetland Sensitivity (SENS)",
  "Wetland Stressors (STR)"
)

# lsheader <- c(
#   "Hydrologic Group (WSg)",
#   "Water Quality (WQg)",
#   "Aquatic Support (ASg)",
#   "Aquatic Habitat (AHg)",
#   "Transition Zone Habitat Group (THg)",
#   "Other Values or Attributes"
# )

# split function and benefit 
classed_df <- classed_df |> 
   dplyr::mutate(service_full_name =  dplyr::case_when(
    indicator  == "AM" ~ "Amphibian Habitat (AM)",
    indicator  == "APP" ~ "Aquatic Primary Productivity (APP)",
    indicator  == "CP" ~ "Carbon Preservation (CP)",
    indicator  == "CRI" ~ "Cultural Recreational Importance (CRI)",
    indicator  == "FH" ~ "Fish Habitat (FH)",
    indicator  == "FR" ~ "Fire Resistance (FR)",
    indicator  == "KMH" ~ "Keystone Mammal Habitat (KMH)",
    indicator  == "NR" ~ "Nitrate Removal and Retention (NR)",
    indicator  == "OE" ~ "Organic Matter Export (OE)",
    indicator  == "PD" ~ "Native Plant Diversity (PD)",
    indicator  == "POL" ~ "Pollinator Habitat (POL)",
    indicator  == "PR" ~ "Phosphorus Retention (PR)",
    indicator  == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
    indicator  == "SENS" ~ "Wetland Sensitivity (SENS)",
    indicator  == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
    indicator  == "SR" ~ "Sediment Retention and stabilization (SR)",
    indicator  == "STR" ~ "Wetland Stressors (STR)",
    indicator  == "WB" ~ "Waterbird Habitat (WB)",
    indicator  == "WS" ~ "Water Storage and Delay (WS)"
  )) 

fntab <- classed_df |> 
  dplyr::filter(service_type == 'f') |> 
  dplyr::mutate(`Function Score (Standardized)` = sta_value,
                `Function Rating` = calibration_scores_summary,
                `Function Score (Raw)` = value,
                `Wetland Functions and other attributes` = service_full_name) |> 
  dplyr::select( `Wetland Functions and other attributes`,
         `Function Score (Standardized)`, `Function Rating`,
         `Function Score (Raw)`)

# split function and benefit 
bntab <- classed_df |> 
  dplyr::filter(service_type == 'b') |> 
  dplyr::mutate(`Benefit Score (Standardized)` = sta_value,
                `Benefit Rating` = calibration_scores_summary,
                `Benefit Score (Raw)` = value,
                `Wetland Functions and other attributes` = service_full_name) |> 
  dplyr::select( `Wetland Functions and other attributes`,
         `Benefit Score (Standardized)`, `Benefit Rating`,
         `Benefit Score (Raw)`)

tball <-  dplyr::full_join(fntab,bntab ) |> 
  dplyr::select("Wetland Functions and other attributes",
                "Function Score (Standardized)", "Function Rating",
                 "Benefit Score (Standardized)",   "Benefit Rating",
                 "Function Score (Raw)","Benefit Score (Raw)")

# add other headers
#headers <- as.data.frame(lsheader ) |> 
#  rename("Wetland Functions and other attributes" = lsheader)

#tball <-  dplyr::bind_rows(tball, headers)

tball <- tball[match(lsorder, tball$"Wetland Functions and other attributes"),]
tball <- tball |> 
  filter(!is.na(`Wetland Functions and other attributes`))

```


```{r, echo = FALSE}
#knitr::kable(tball)

```


```{r generate table, echo = FALSE}
gt_tbl <- gt(tball)

gt_tbl <-
  tball |>
  gt(rowname_col = "Wetland Functions and other attributes") |> 
  tab_stubhead(label = "Wetland Functions and other attributes") #|> 
# tab_header(
#    title = "Wetland Function and Benefits scores and ratings",
#  )
gt_tbl <-
  gt_tbl |>
  tab_row_group(
    label = "Other Values or Attributes",
    rows = 15:19
)|>
  tab_row_group(
    label = "Transition Zone Habitat Group (THg)",
    rows = 11:14
  )|> 
  tab_row_group(
    label = "Aquatic Habitat (AHg)",
    rows = 8:10
  )|> 
  tab_row_group(
    label = "Aquatic Support (ASg)",
    rows = 5:7
  ) |>  
  tab_row_group(
    label = "Water Quality (WQg)",
    rows = 2:4 
  ) |>
  tab_row_group(
    label = "Hydrologic Group (WSg)",
    rows = 1
  ) 

#gt_tbl

gt_tbl <- gt_tbl |> 
  tab_spanner(
    label = "Function",
    columns = c(`Function Score (Standardized)`, `Function Rating`)
  )|>
  tab_spanner(
    label = "Benefit",
    columns = c("Benefit Score (Standardized)", "Benefit Rating")
  ) |> 
  tab_spanner(
    label = "Raw Scores",
    columns = c( "Function Score (Raw)","Benefit Score (Raw)")
  ) |> 
  cols_align(align = "center") |> 
  cols_align(align = "left", columns = "Wetland Functions and other attributes") |> 
#   tab_source_note(
#     source_note = md("NOTE: A score of 0 does not always mean the function or value is absent from the wetland. It usually
# means that this wetland has equal or less capacity than the lowest-scoring one, for that function or
# value, from among the 100+ calibration wetlands that were assessed previously for the EcoProvince.")) |>   
  tab_options(
    table.font.size = "small"
  )

gt_tbl
      

```

<br>


NOTE: A score of 0 does not always mean the function or value is absent from the wetland. It usually
means that this wetland has equal or less capacity than the lowest-scoring one, for that function or
value, from among the 100+ calibration wetlands that were assessed previously for the EcoProvince.

<br>

#### Output Description

The **Score** columns present the numeric score of a function, value, or attribute after it has been
mathematically adjusted (standardized) to a full 0-10 point scale, based on calibration data from 100+
test sites.

The **Rating** columns indicate which of three rating categories [Lower (L), Moderate (M), Higher(H)]
each normalized score is assigned to. Ratings convey the relative meaning of the numeric score and
allow for comparison across different functions and benefits. The breaks between ratings differ for
each function and benefit, based on the distribution of scores from the 100+ calibration wetlands.
<br>


```{r raw_plots, echo = FALSE, width = 10}
#raw_plot

```

<br>
<br>


### 2.2 Ecosystem Scores and Ratings

```{r, echo = FALSE}
# # generate raw values plots based on themes 
# 
# # functions 
# 
# ff <- classed_df |> 
#   mutate(service_name = indicator) |> 
#   mutate(service_full_name = case_when(
#     service_name == "AM" ~ "Amphibian Habitat (AM)",
#     service_name == "APP" ~ "Aquatic Primary Productivity (APP)",
#     service_name == "CP" ~ "Carbon Preservation (CP)",
#     service_name == "CRI" ~ "Cultural Recreational Importance (CRI)",
#     service_name == "FH" ~ "Fish Habitat (FH)",
#     service_name == "FR" ~ "Fire Resistance (FR)",
#     service_name == "KMH" ~ "Keystone Mammal Habitat (KMH)",
#     service_name == "NR" ~ "Nitrate Removal and Retention (NR)",
#     service_name == "OE" ~ "Organic Matter Export (OE)",
#     service_name == "PD" ~ "Native Plant Diversity (PD)",
#     service_name == "POL" ~ "Pollinator Habitat (POL)",
#     service_name == "PR" ~ "Phosphorus Retention (PR)",
#     service_name == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
#     service_name == "SENS" ~ "Wetland Sensitivity (SENS)",
#     service_name == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
#     service_name == "SR" ~ "Sediment Retention and stabilization (SR)",
#     service_name == "STR" ~ "Wetland Stressors (STR)",
#     service_name == "WB" ~ "Waterbird Habitat (WB)",
#     service_name == "WS" ~ "Water Storage and Delay (WS)"
#   )) |> 
#   select( service_type, value, service_name, service_full_name) |> 
#   mutate(service_type_full = case_when(
#     service_type == "f" ~ "function",
#     service_type == "b" ~ "benefit"
#   ))
# 
# ggplot(ff, aes(x = service_type_full, y = service_full_name)) +
#   geom_tile(aes(fill = value)) +
#   geom_text(aes(label = round(value, 1))) +
#   #scale_fill_continuous("blue")
#   theme(palette.colour.continuous = c("#DEEBF7", "#9ECAE1", "#3182BD"))+
#   labs(
#     x = "Ecosystem Score",
#     y = "Ecosystem Function"
#   )

```



```{r build overall plots, echo=FALSE, fig.width=8, fig.height=7}
tt <- calibration_scores_summary |>
   dplyr::group_by(.data$service) |>
   dplyr::select(-n) |>
  tidyr::pivot_wider(names_from = jenks, values_from = c(min, max)) |>
  dplyr::mutate(
    H = 10.1 - min_H,
    M = min_H - min_M,
    L = min_M
  )

ttt <- tt |>
  tidyr::pivot_longer(
    cols = c(H, M, L, -min_H, -min_L, -min_M, -max_H, -max_L, -max_M),
    names_to = "catergory",
    values_to = "threshold"
  ) |>
  dplyr::select(-c(min_H, min_L, min_M, max_H, max_L, max_M))

ttt <- ttt |>
  dplyr::mutate(class = ordered(catergory, levels = c("H", "M", "L"))) |>
  dplyr::mutate(service_full_name = case_when(
    service_name == "STR" ~ "Wetland Stressors (STR)",
    service_name == "SENS" ~ "Wetland Sensitivity (SENS)",
    service_name == "CRI" ~ "Cultural Recreational Importance (CRI)",
    service_name == "FR" ~ "Fire Resistance (FR)",
    service_name == "CP" ~ "Carbon Preservation (CP)",
    service_name == "POL" ~ "Pollinator Habitat (POL)",
    service_name == "PD" ~ "Native Plant Diversity (PD)",
    service_name == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
    service_name == "KMH" ~ "Keystone Mammal Habitat (KMH)",
    service_name == "AM" ~ "Amphibian Habitat (AM)",
    service_name == "WB" ~ "Waterbird Habitat (WB)",
    service_name == "FH" ~ "Fish Habitat (FH)",
    service_name == "APP" ~ "Aquatic Primary Productivity (APP)",
    service_name == "OE" ~ "Organic Matter Export (OE)",
    service_name == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
    service_name == "NR" ~ "Nitrate Removal and Retention (NR)",
    service_name == "PR" ~ "Phosphorus Retention (PR)",
    service_name == "SR" ~ "Sediment Retention and stabilization (SR)",
    service_name == "WS" ~ "Water Storage and Delay (WS)"
  )) |> 
   dplyr::mutate(service_full_name = ordered(service_full_name, 
                                             levels = c( 
     "Wetland Stressors (STR)",
   "Wetland Sensitivity (SENS)",
    "Cultural Recreational Importance (CRI)",
   "Fire Resistance (FR)",
    "Carbon Preservation (CP)",
   "Pollinator Habitat (POL)",
   "Native Plant Diversity (PD)",
   "Raptor and Wetland Songbird Habitat (RSB)",
    "Keystone Mammal Habitat (KMH)",
    "Amphibian Habitat (AM)",
    "Waterbird Habitat (WB)",
  "Fish Habitat (FH)",
    "Aquatic Primary Productivity (APP)",
   "Organic Matter Export (OE)",
     "Stream Flow and Temperature Support (SFTS)",
  "Nitrate Removal and Retention (NR)",
     "Phosphorus Retention (PR)",
    "Sediment Retention and stabilization (SR)",
     "Water Storage and Delay (WS)")))


# match format for site data
classed_df <- classed_df |>
  dplyr::mutate(
    service = paste0(indicator, "_", service_type),
    service_name = indicator,
    threshold = value
  ) |>
  dplyr::mutate(service_full_name = case_when(
    service_name == "STR" ~ "Wetland Stressors (STR)",
    service_name == "SENS" ~ "Wetland Sensitivity (SENS)",
    service_name == "CRI" ~ "Cultural Recreational Importance (CRI)",
    service_name == "FR" ~ "Fire Resistance (FR)",
    service_name == "CP" ~ "Carbon Preservation (CP)",
    service_name == "POL" ~ "Pollinator Habitat (POL)",
    service_name == "PD" ~ "Native Plant Diversity (PD)",
    service_name == "RSB" ~ "Raptor and Wetland Songbird Habitat (RSB)",
    service_name == "KMH" ~ "Keystone Mammal Habitat (KMH)",
    service_name == "AM" ~ "Amphibian Habitat (AM)",
    service_name == "WB" ~ "Waterbird Habitat (WB)",
    service_name == "FH" ~ "Fish Habitat (FH)",
    service_name == "APP" ~ "Aquatic Primary Productivity (APP)",
    service_name == "OE" ~ "Organic Matter Export (OE)",
    service_name == "SFTS" ~ "Stream Flow and Temperature Support (SFTS)",
    service_name == "NR" ~ "Nitrate Removal and Retention (NR)",
    service_name == "PR" ~ "Phosphorus Retention (PR)",
    service_name == "SR" ~ "Sediment Retention and stabilization (SR)",
    service_name == "WS" ~ "Water Storage and Delay (WS)"

  ))


# summary of the ranks overall
all_ranks_f <- classed_df |>
  dplyr::filter(service_type == "f") |>
  dplyr::group_by(calibration_scores_summary) |>
  dplyr::summarise(n = n())

# functions
ft <- ttt |>
  dplyr::filter(service_type == "f")

fclass <- classed_df |>
  dplyr::filter(service_type == "f") |>
   dplyr::filter(!is.na(value)) |>
   dplyr::left_join(outsum)

fn_plot <- ggplot2::ggplot(ft, aes(x = threshold, y = service_full_name)) +
  ggplot2::geom_bar(aes(fill = class), stat = "identity", position = "stack") +
  scale_fill_viridis_d() +
  # geom_text(data= fclass,aes(x= max,y=service_full_name,label=n), color = "darkgrey",vjust=0) +
  ggplot2::geom_point(data = fclass, aes(x = threshold, y = service_full_name, size = 1.25), colour = "darkgrey") +
  labs(
    x = "Ecosystem Score",
    y = "Ecosystem Function",
    fill = "Rating"
  ) +
  scale_size(guide = 'none')+
  theme_minimal()


```


The plots below shows where `r sname` scores relative to the baseline values for the `r ecop` Ecoprovince. The grey dot indicates the score for the single site, while the yellow, aqua and purple bars show the threshold values for Lower, Medium and Higher ranked sites within the Ecoprovince.  

```{r add plot, echo = FALSE, fig.width=8, fig.height=6, units = cm}
 fn_plot
```

```{r build benefit plots, echo = FALSE, fig.width=8, fig.height=6}
 # Benefit
 bt <- ttt |> 
   dplyr::filter(service_type == "b") 
 
 bclass <-  classed_df|> 
   dplyr::filter(service_type == "b")|> 
   filter(!is.na(value))
 
 
bn_plot <-  ggplot(bt , aes(x = threshold,y = service_full_name  )) +
    geom_bar(aes(fill = class), stat = "identity", position = "stack")+
    scale_fill_viridis_d()+ 
   #  geom_text(data= dfTab,aes(x=x,y=Freq,label=lab),vjust=0) +
    geom_point(data =  bclass, aes( x = threshold, y = service_full_name, size = 1.25 ), colour = "darkgrey")+
   labs(
        x = "Ecosystem Score",
        y = "Ecosystem Benefit",
    fill = "Rating") +
   scale_size(guide = 'none')+
   theme_minimal()

#bn_minplot <- bn_plot +
#   theme(axis.text.y=element_blank())

# 
#  #summary of the ranks overall 
#  all_ranks_b <- bclass |> 
#   group_by(calibration_scores_summary) |>
#   summarise(n = n())
   
```


```{r add benefit plot, echo = FALSE, fig.width=8, fig.height=6 ,units = cm}
 bn_plot 
```


