---
title: "Advanced-topics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced-topics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(wespr)
```


This vignette walks through some helper functions for those interested in more advanced topics relating to the wespr package. **This is not required to calculate scores** and is intended for advanced R users. 



## 1) Updating internal data

### Weighs and ranking tables

The wespr package contains two internal datasets which facilitate question validation and indicator calculations: `question_metadata` and `indicator_weightings`.

These are updated by:

1. Update the Google Sheet at:  https://docs.google.com/spreadsheets/d/1l2h7Z65H5z0cKv_gvorxkT6k9LC7ZKLS/edit#gid=924841838 (you need to be authorized to edit this document)

2. Run the script in `data-raw/internal-data.R`

3. Reload/reinstall the package.

4. Run `devtools::check()` to make sure the changes you made to the Google Sheet haven't broken any existing behaviour.

5. Commit the changes to `R/sysdata.rda` (this file contains both datasets)




### Calibration data 

The wespr package contains the summarized `calibration_scores`. These will require updating as more calibration data becomes available. 


These are updated by:

1. Run the script in `data-raw/upload-calibration-data.R`

3. Reload/reinstall the package.

4. Run `devtools::check()` to make sure the changes you made to the Google Sheet haven't broken any existing behaviour.

5. The calibration data is accessible to users via the `calibration_scores` function.



```{r extra notes, eval= FALSE,include = FALSE}
## 1) Calculating Jenks Scores for a new Ecoprovince

Another method to step through the entire process of calculating site scores, and assigning them to a high medium and low category is to use the `calculate_jenks_score()` function. This function requires the data to be loaded using the `load_wesp_data()` function as above. 

This will calculate the indicator scores for each site, normalize the data across all sites for the given ecosystem function or benefit then calculate the Jenks breaks for each value to assign high, medium and low categories. The process will also flag unusual results, such as all values being the same across all sites. 

Note the `calculate_jenks_score()` function includes the steps described above to calculate all site scores. An example workflow using this function would be: 


#```{r,eval = FALSE}
wesp_file <- system.file("input_data/reference_singlesite.csv", package = "wespr")
wesp_data <- load_wesp_data(wesp_file)
calculate_jenks_score(wesp_data, out_dir = "temp",  out_name = "wesp_scores.csv")

```



## 2) Anatomy of the `wesp_site` object

As mentioned above, this is a large, complex object, and it is almost always 
best to interact with the `wesp_site` object via the provided functions. If 
there is functionality that is missing please 
[open an issue](https://github.com/BCWF-Wetlands/wespr/issues). 

It has four elements:

```{r,eval = FALSE}
length(site)
names(site)
```

The first three (`site_name`, `questions`, and `derived_values`) are populated when 
the `wesp_site` object is created, but the `indicators` element is just a placeholder, until `calc_indicators()` has been run:

```{r,eval = FALSE}
site$site_name

# The questions element is large and complex, so we will just show the names:
names(site$questions)

site$derived_values

site$indicators
```



