---
title: "beginner_workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{beginner_workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wespr)
```

To help with the use of wespr, we have developed a workflow to streamline the steps to generate wetland ecosystem values. 

The follow function steps through from raw files to calculate wespr scores and assign jenks breaks. 
Users need to define the following parameters: 

- field_data : file path to location of field survey123 output (.csv)
- desktop_data : file path to location of desktop survey123 output (.csv)
- Eco : ecoprovince code (i.e "GD")
- out_dir : file path to location where output files will be saved


```{r, eval = FALSE}
# Define the location of the survey123 Desktop data 
field_data <- system.file(file.path('extdata','WESP_FIELDV1.csv'), package = "wespr")
desktop_data <- system.file(file.path('extdata','WESP_DESKTOPV1.csv'), package = "wespr")

out <- wespr_workflow(field_data = field_data,
                      desktop_data = desktop_data, 
                      EcoP = "GD",
                      out_dir = "temp")

write.csv(out, fs::path("temp", "wespr_outputs.csv"))

```
