---
title: "calculate_wespr_single_site"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{calculate_wespr_single_site}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup,include=FALSE}
library(wespr)
```


## 1) Import data in wesp site object

Once the wespr package has been installed, and the raw data processed using the steps outlined in the previous [vignette](https://bcwf-wetlands.github.io/wespr/articles/prepare-rawdata.html), we can calculate the wespr scores and compare the scores to the ecoprovince calibration data. 


An example file in the package, which we can use for demonstration purposes:

```{r, eval = FALSE}
# point to example file
wesp_file <- system.file("input_data/reference_singlesite.csv", package = "wespr")

# load the wesp data 
wesp_data <- load_wesp_data(wesp_path)

# we can take a look at this file to see the format
head(wesp_data)

```

```
q_no, response_no,site_1
F1,1,1
F1,2,0
F1,3,1
F1,4,1
F1,5,1
F1,6,1
F2,A1,0
F2,A2,0
F2,B1,0
F2,B2,1
```

We will also generate a key for the site names which we can use later after processing the site information. 

```{r, eval = FALSE}
# generate a key for site names 
wespkey <- wespr::generate_ids(wesp_data)

```


Next we convert the data into a special `wesp_site` object. This process validates
the input data, and will give informative errors if a question is not answered correctly.
It also calculates many "derived values" from the questions, which are common inputs into different
indicator calculations.

If there is more than one site in the input data, you can select which site you
would like to use --- if no site is selected it will default to choosing the first 
site in the file. See [here](https://bcwf-wetlands.github.io/wespr/articles/calculate_wespr_multi_site.html) for a step by step guide to processing multisite datasets.

```{r, eval = FALSE}
site <- as.wesp_site(wesp_data, site = 1)

```

You should not normally need to work with the internals of the `wesp_site` object
as it is quite complex, however we can get an overview of what is in it by just 
typing the name of the object:

```{r, eval = FALSE}
site

```

## 2) Explore in wesp site object

We can see the structure of the wesp_site object includes a site name, any warnings or flags for missing data and a list of derived values.You can see the indicators have not yet
been calculated at this stage.

```
A wesp_site object

Site:  site_1 

Incomplete Questions:  F59 
  * Please ensure that it is valid to leave these questions unanswered.

Derived values:
  *  AllWater = 0
  *  NeverWater = 0
  *  NoSeasonal = 0
  *  NoPersis = 0
  *  TempWet = 0
  *  AllPermW = 0
  *  HiFlucW = 0
  *  TooShallow = 0
  *  NoPond = 0
  *  NoDeepPonded = 0
  *  NoOW = 0
  *  NoOutletX = 0
  *  NoOutlet = 0
  *  Inflow = 1
  *  Disturb = 1
  *  FishFound = 0
  *  Moose = 0
  *  Beaver = 1
  *  Muskrat = 0
  *  Bear = 1
  *  Caribou = 0
  *  NoCA = 0
  *  Fishless = 0
  *  GDeco = 0
  *  CMeco = 0
  *  SIMeco = 1
  *  BPeco = 0
  *  TPeco = 0
  *  OutMap = 1
  *  S1_sum = 4
  *  S1_subscore = 0.33
  *  S2_sum = 1
  *  S2_subscore = 0.11
  *  S3_sum = 1
  *  S3_subscore = 0.11
  *  S4_sum = 4
  *  S4_subscore = 0.33
  *  S5_sum = 0
  *  S5_subscore = 0
  *  S6_sum = 2
  *  S6_subscore = 0.33

Indicators:
All indicators are NULL. Run `calc_indicators()` to calculate them.

```
## 3) Calculate indicator raw scores

Now, we can calculate the indicator scores with the `calc_indicators()` function.

You must assign the output of this function back to the original `wesp_site` 
object, as it updates the `indicators` with the calculated values:

```{r, eval = FALSE}
site <- calc_indicators(site)
```

If we view the site object again, we can see the calculated indicator scores.

```{r, eval = FALSE}
site

```
```
A wesp_site object

Site:  site_1 

Incomplete Questions:  F59 
  * Please ensure that it is valid to leave these questions unanswered.

Derived values:
  *  AllWater = 0
  *  NeverWater = 0
  *  NoSeasonal = 0
  *  NoPersis = 0
  *  TempWet = 0
  *  AllPermW = 0
  *  HiFlucW = 0
  *  TooShallow = 0
  *  NoPond = 0
  *  NoDeepPonded = 0
  *  NoOW = 0
  *  NoOutletX = 0
  *  NoOutlet = 0
  *  Inflow = 1
  *  Disturb = 1
  *  FishFound = 0
  *  Moose = 0
  *  Beaver = 1
  *  Muskrat = 0
  *  Bear = 1
  *  Caribou = 0
  *  NoCA = 0
  *  Fishless = 0
  *  GDeco = 0
  *  CMeco = 0
  *  SIMeco = 1
  *  BPeco = 0
  *  TPeco = 0
  *  OutMap = 1
  *  S1_sum = 4
  *  S1_subscore = 0.33
  *  S2_sum = 1
  *  S2_subscore = 0.11
  *  S3_sum = 1
  *  S3_subscore = 0.11
  *  S4_sum = 4
  *  S4_subscore = 0.33
  *  S5_sum = 0
  *  S5_subscore = 0
  *  S6_sum = 2
  *  S6_subscore = 0.33
  
  Indicators:
  * WS:  
    - fun: 2.97 
    - ben: 3.3 
  * SR:  
    - fun: 0 
    - ben: 4.09 
  * PR:  
    - fun: 4.77 
    - ben: 3.03 
  * CP:  
    - fun: 7.06 
  * FR:  
    - fun: 4.93 
    - ben: 1.06 
  * SENS:  
    - fun: 7.02 
  * STR:  
    - fun: 6.61 
  * NR:  
    - fun: 5.27 
    - ben: 3.57 
  * APP:  
    - fun: 3.82 
    - ben: 4.39 
   * PD:  
    - fun: 5.59 
    - ben: 4.19 
  * KMH:  
    - fun: 6.41 
    - ben: 3.93 
  * WB:  
    - fun: 5.62 
    - ben: 10 
  * POL:  
    - fun: 5.95 
    - ben: 2.25 
  * RSB:  
    - fun: 5.05 
    - ben: 6.11 
  * OE:  
    - fun: 4.1 
  * AM:  
    - fun: 3.81 
    - ben: 8.1 
  * FH:  
    - fun: 3.08 
    - ben: 3.72 
  * SFTS:  
    - fun: 3.46 
    - ben: 3.63 
  * CRI:  
    - ben: 4.93 

* Retrieve indicator scores with `get_indicator_scores()`
```

We probably want to get the indicator scores out as a usable object. We can do 
that with `get_indicator_scores()`, which gives them to us as a `data.frame`:

```{r, eval = FALSE}
ind_scores <- get_indicator_scores(site)
```

```
  site   indicator    fun   ben
   <chr>  <chr>      <dbl> <dbl>
 1 site_1 WS         3.61   6.98
 2 site_1 SR         0.778  4.28
 3 site_1 PR         6.20   3.06
 4 site_1 CP         6.78  NA   
 5 site_1 FR         1.67   7.51
 6 site_1 SENS       4.86  NA   
 7 site_1 STR        9.58  NA   
 8 site_1 NR         5.52   3.06
 9 site_1 APP        0      2.99
10 site_1 PD         0      5.51
11 site_1 KMH        2.99   5.60
12 site_1 WB         0     10   
13 site_1 POL        6.30   1.67
14 site_1 RSB        6.94   4.83
15 site_1 OE         4.54  NA   
16 site_1 AM         5.03   3.10
17 site_1 FH         0      3.33
18 site_1 SFTS       2.94   3.27
19 site_1 CRI       NA      6.58

```
An example format of the ind_score output is shown above, with the service type (rows) for each site with a column for function (fun) and benefit (ben). Note not every service has both a function and benefit so you will see NA values.  

```{r see ind scores, eval = FALSE}
# add site specific names
ind_scores <- dplyr::left_join(wespkey, ind_scores, by = "site")
 
# lets see what it looks like now
ind_scores

```


There are a number of helper functions to extract information from the wespr_site object. For example we can gather just the responses with `get_responses()`:

```{r, eval = FALSE}
get_responses(site)
```
```
no  question                            response_no   value
F1	Vegetation Height & Form Diversity	  F1_1	        3.00	
F1	Vegetation Height & Form Diversity	  F1_2	        0.00	
F1	Vegetation Height & Form Diversity	  F1_3	        2.00	
F1	Vegetation Height & Form Diversity	  F1_4	        2.00	
F1	Vegetation Height & Form Diversity	  F1_5	        1.00	
F1	Vegetation Height & Form Diversity	  F1_6	        3.00	
F10	Dense Moss Extent	                  F10_1	        0.00	
F10	Dense Moss Extent	                  F10_2	        1.00	
F10	Dense Moss Extent	                  F10_3       	0.00	
F10	Dense Moss Extent	                  F10_4	        0.00	
``` 

We can also get out a data.frame of derived values, those values which are 
calculated from the responses, and used as inputs into many indicators.

```{r,eval = FALSE}
get_derived_values(site)
```
```
name          value

AllWater	    0.0000000			
NeverWater	    0.0000000			
NoSeasonal	    0.0000000			
NoPersis	    0.0000000			
TempWet	        0.0000000			
AllPermW	    0.0000000			
HiFlucW	        0.0000000			
TooShallow	    0.0000000			
NoPond	        0.0000000			
NoDeepPonded	0.0000000	
```

## 4) Assign Jenks scores

Once we have calculated our scores we may wish to compare to see how our particular site compares with other wetlands in the same Ecoprovince. [Wespr protocol](https://bcwfwatershedteam.ca/wetland-ecosystem-services-protocol/) is based on wetland assessment in eight Ecoprovinces (Georgia Depression (GD), Southern Interior (SI),
Central Interior (CI), Southern Interior Mountains (SIM), Sub-Boreal Interior (SBI), Boreal and Taiga Plains (BTP), Northern Boreal Mountains (NBM), and Coast and Mountains (CM)).

Using our data from above we see the `ind_score` object is a data table with a row for each indicator and columns for the raw values for functions (fun) and benefits (ben).  An example format of the ind_score output is shown, with the service type (rows) for each site with a column for function (fun) and benefit (ben). Note not every service has both a function and benefit so you will see NA values.  


```{r, eval = FALSE}
ind_scores

```
```
site    wetland_id indicator fun       ben

site_1	SIM_90009	  WS	    2.968247	3.304924
site_1	SIM_90009	  SR	    0.000000	4.094300
site_1	SIM_90009	  PR	    4.772264	3.031987
site_1	SIM_90009	  CP	    7.055459	NA
site_1	SIM_90009	  FR	    4.930556	1.056566
site_1	SIM_90009	  SENS	    7.015852	NA
site_1	SIM_90009	  STR	    6.607143	NA

```


We use the `assign_jenks_score()` function to assign a catergory of Low/Medium/High to each of the values, based on where our site sits in relation to 100 or more calibration sites for the given ecoprovince. 

The calibration data is stored within the package and can be [updated by package maintainers](https://bcwf-wetlands.github.io/wespr/articles/advanced-topics.html). This is an on-going project and data will be added as comes available. Steps to update the calibration data are outline [here](https://github.com/BCWF-Wetlands/wespr/tree/main/data-raw). Note this is an advanced/admin level action and not required to calculate values. Please contact the package dev team for assistance. 

Threshold values to define the low , medium or high class are determined by normalising all the reference data for each value and then looking for natural breask in the distribution (i.e. Jenks Breaks). These thresholds are then used as benchmarks and compared to our site of interest to determine which category it falls within. 

Note in some cases, out site value may fall below or above data gathers for the calibration sites. In this case, a warning message is provided to the user.


To run this function we require: 

- the indicator score data (i.e. ind_scores in this workflow), 
- calibration data (this is stored as a file within the wespr package)
- name of Ecoprovince in which the site occurs.
- if a report is to be generated (T/F)
- if a report = TRUE, an output directory where the report is to be saved. 


If the ecoprovince does not yet have calibration data loaded, users should select the next best ecoprovince available. 

```{r eval = FALSE}
out <- assign_jenks_score(
  ind_scores, 
  calibration_scores, 
  EcoP = "GD", 
  report = FALSE, 
  output_dir = "temp")

```

The output scores will include a column with a calibration_score (L,M,H)

```
site   indicator value service_type calibrated_score
   <chr>  <chr>     <dbl> <chr>        <chr>           
 1 site_1 WS        3.61  f            L               
 2 site_1 SR        0.778 f            M               
 3 site_1 PR        6.20  f            M               
 4 site_1 CP        6.78  f            L               
 5 site_1 FR        1.67  f            L               
 6 site_1 SENS      4.86  f            L               
 7 site_1 STR       9.58  f            H               
 8 site_1 NR        5.52  f            M               
 9 site_1 APP       0     f            L               
10 site_1 PD        0     f            L  

```

Now you have a data table with all indicators, service_type (function or benefit), raw scores and calibrated score (L/M/H). THis can be saved or exported as a .csv to be further explored. 
