---
title: "r2 Simple Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{r2 Simple Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wespr)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To help with the use of WESPR-BC, we have developed a workflow to streamline the steps to generate wetland ecosystem values. 

The follow function steps through from raw files to calculate WESPR-BC scores, assign jenks breaks and output a report.
Users need to define the following parameters: 

- field_data : file path to location of field survey123 output (.csv)
- desktop_data : file path to location of desktop survey123 output (.csv)
- Eco : BC Ecoprovince code (i.e "GD")
- out_dir : file path to location where output files will be saved

```{r, eval = FALSE}
# Define the location of the survey123 Desktop data 
field_data <- system.file(file.path('extdata','WESP_FIELDV1.csv'), package = "wespr")
desktop_data <- system.file(file.path('extdata','WESP_DESKTOPV1.csv'), package = "wespr")

out <- wespr_workflow(field_data = field_data,
                      desktop_data = desktop_data, 
                      EcoP = "GD",
                      out_dir = "temp", 
                      report_dir = "temp")

write.csv(out, fs::path("temp", "wespr_outputs.csv"))

```
